---
title: "lab4"
author: "Ayimen H."
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Left-handedness

(a) Suppose I randomly pick 200 people. What is the chance I will see fewer than 
15 people who are left-handed?
``` {r}
p_left_handed <- 0.10
n_a <- 200
x_a <- 14
prob_a <- pbinom(x_a, n_a, p_left_handed)

prob_a
```
(b) Suppose I randomly pick 300 people. What is the chance I will see at least 
40 people who are left-handed?
```{r}
n_b <- 300
x_b <- 39
prob_b <- 1 - pbinom(x_b, n_b, p_left_handed)

prob_b
```
(c) Suppose that the Seattle Mariners have 21 pitchers. Of those 21 pitchers, 4 
are left-handed. If we assume the probability of a pitcher being left-handed is 
the same as the probability of any randomly selected person from the general 
population being left-handed, what would be the probability of seeing at least 4 
left-handed pitchers out of 21?
```{r}
n_c <- 21
x_c <- 3
prob_c <- 1 - pbinom(x_c, n_c, p_left_handed)

prob_c
```
(d) Based on your answer from part (c), does our assumption about the probability 
of a pitcher being left-handed seem reasonable? If not, how does the probability 
differ from the probability for the general population?

Based on the 15% from the previous question I believe the assumption that 10% of 
pitchers being left-handed is reasonable

(e) The Seattle Mariners have 3 catchers. All 3 of them are right-handed. If we 
assume the probability of a catcher being left-handed is the same as the 
probability of any randomly 2 selected person from the general population being 
left-handed, what would be the probability of seeing no left-handed catchers out 
of 3?
```{r}
n_e <- 3
x_e <- 0
prob_e <- dbinom(x_e, n_e, p_left_handed)

prob_e
```
(f) Based on your answer from part (e), does our assumption about the probability 
of a catcher being left-handed seem reasonable? If not, how does the probability 
differ from the probability for the general population?

Based on e I believe it doesn't differ too much considering the sample size being
so small. The 72.9% of no lefties reinforces our assumption.

(g) Suppose we were to look at a larger sample of catchers. If every catcher we 
sampled were right-handed, how many catchers would we need to sample before you 
would conclude that the true probability of a catcher being left-handed is less 
than 10%?

```{r}
# couldn't figure out how to do this with a distribution

p_left_handed <- 0.10
n_g_alt <- ceiling(log(0.05) / log(1 - p_left_handed))

n_g_alt
```

2. Computer server

(a) What is the chance that more than 15 students will log into the server in a 
particular hour?
```{r}
lambda_hourly <- 10
x_2a <- 15
prob_2a <- 1 - ppois(x_2a, lambda_hourly)

prob_2a
```
(b) What is the chance of seeing exactly 10 students log into the server in a 
particular hour?
```{r}
x_2b <- 10
prob_2b <- dpois(x_2b, lambda_hourly)

prob_2b
```
(c) What is the chance of fewer than 15 students logging into the server in a 
two-hour period?
```{r}
lambda_two_hours <- 2 * lambda_hourly
x_2c <- 14
prob_2c <- ppois(x_2c, lambda_two_hours)

prob_2c
```
(d) In designing the server, you must decide the maximum number of students that 
it can accommodate at one time. The more students you allow it to accommodate, 
the more expensive it will be. But if more students attempt to log in during a 
single hour than it can accommodate, it will crash. How many students should you 
design it to accommodate if you want there to be at most a 1% chance that it will 
crash during any particular hour?
```{r}
lambda_hourly <- 10
max_students <- qpois(0.99, lambda_hourly)

max_students
```
